---
title: "Classification with geomedian"
author: "Michael Wellington"
date: "2024-07-16"
output: html_document
---

# Load packages

```{r load, include=TRUE, warning=FALSE}
library(sf)
library(terra)
library(sits)
library(rstac)
library(rnaturalearth)
library(tidyverse)
```
# Load and plot Murang'a boundary

We'd like to limit our classification to the Murang'a county in Kenya. We've obtained the shapefile boundary from the [Database of Global Administrative Areas (GADM) website](https://gadm.org/) and stored it in this repositories Data folder.

```{r muranga bound}
kenya_a2 <- read_sf('data/gadm41_KEN_2.shp')
kenya_muranga <- kenya_a2 %>% filter(NAME_1 == "Murang'a")
plot(kenya_muranga['NAME_2'])
```
# Classes
* Water
* Grassland
* Cropland
* Urban
* Forest


# Load geomedian

This script loads a geomedian, a product unique to DE Africa because it provides a cloud-free composite over a given period.

SITS accepts a bounding box (bbox) as an area of interest, so we use the spatial extent of the Murang'a County for this parameter.

```{r rstacload}
dea_gm_cube <- sits_cube(
  source = "DEAFRICA",
  collection = "GM-LS8-LS9-ANNUAL",
  roi = c(
    lon_min = ext(kenya_muranga)$xmin[[1]], lat_min = ext(kenya_muranga)$ymin[[1]],
    lon_max = ext(kenya_muranga)$xmax[[1]], lat_max = ext(kenya_muranga)$ymax[[1]]
  ),
  bands = c("SR_B2", "SR_B3", "SR_B4", "SR_B5", "SR_B6", "SR_B7", "SMAD"),
  start_date = "2019-01-01",
  end_date = "2022-12-31"
)
plot(dea_gm_cube, red = "B04", blue = "B02", green = "B03")
```
```{r trainingdat}
muranga_points <- st_read('data/Muranga_classes.shp')
muranga_points
```

```{r training_dat}

series <- sits_get_data(
    cube=dea_gm_cube,
    label_attr = "Class",
    samples='data/Muranga_classes.shp')
```

```{r series}
summary(series)
```
```{r bands}
dea_gm_cube$file_info
```

```{r sampleplot}
series %>% 
  sits_apply(NDVI = (B05 - B04) / (B05 + B04))  %>% 
  sits_apply(EVI = 2.5 * (B05 - B04) / ((B05 + 6.0 * B04 - 7.5 * B02) + 1.0))  %>% 
  sits_select(bands = c("NDVI", "EVI"))

```
```{r modelsetup}

set.seed(20)

rfor_model <- sits_train(
  series,
  ml_method = sits_rfor()
)
plot(rfor_model)
```
```{r randfor}
muranga_probs <- sits_classify(
  data = dea_gm_cube,
  ml_model = rfor_model,
  output_dir = "data",
  version = "rf-raster",
  multicores = 4,
  memsize = 16
)

plot(muranga_probs, labels = "Forest", palette = "YlGn")
```
## Classified Map
```{r class map}
muranga_smooth <- sits_smooth(
  cube = muranga_probs,
  multicores = 4,
  memsize = 12,
  output_dir = "data",
  version = "rf-raster"
)
```


```{r class map}
muranga_class <- sits_label_classification(
  cube = muranga_smooth,
  multicores = 4,
  memsize = 12,
  output_dir = "data",
  version = "rf-raster"
)

# Plot the thematic map
plot(muranga_class, tile=muranga_class$tile[2])
```
## Create  mosaic of classified classes
```{r mosaic}
muranga_mosaic <- sits_mosaic(
  roi = st_transform(st_union(kenya_muranga), "EPSG:6933"),
  crs = "EPSG:6933",
  cube = muranga_class,
  multicores = 4,
  output_dir = "data",
  version = "rf-raster"
)

plot(muranga_mosaic)
```

# Conclusion

We can see that DE Africa's unique geomedian is accessible for classification in R.
