---
title: "Data Cubes"
author: "Michael Wellington"
date: "2024-07-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Earth Observation Data Cubes

The SITS package implements a concept of data cubes to store and deal with data in space and time. This is not to be confused with the Open Data Cube, which is a broader way of organising Earth Observation data on which Digital Earth Africa is implemented.

In the last script, we loaded Digital Earth Africa Sentinel-2 data using the `sits_cube()` function which returned a data cube, stored locally in our R environment. This script will further explore data cubes objects and conducting operations on them.

To begin, packages are loaded.
```{r packages, include=TRUE}
library(sf)
library(terra)
library(sits)
library(rstac)
library(parallel)
library(tidyverse)
```

## Load a Sentinel-2 data cube

Sentinel-2 images for an area around the Kenyan capital, Nairobi, is loaded and stored as a data cube object. We can plot the first image in the series, noting that it is cloudy.

```{r s2cube, echo=TRUE}
dea_s2_cube <- sits_cube(
  source = "DEAFRICA",
  collection = "SENTINEL-2-L2A",
  roi = c(
    lon_min = 36.65, lat_min = -1.48,
    lon_max = 36.95, lat_max = -1.18
  ),
  bands = c(
    "B02", "B03", "B04",
    "B05", "B06", "B07",
    "B08", "B8A", "B11",
    "B12", "CLOUD"
  ),
  start_date = "2023-01-01",
  end_date = "2023-01-31"
)
plot(dea_s2_cube, red = "B04", blue = "B02", green = "B08")
```

Inspecting the data cube object, below, shows that two tiles have been returned. We can check the tiles to see their extent. 

```{r object, echo=TRUE}
dea_s2_cube
```
```{r plottile1, echo=TRUE}
plot(dea_s2_cube %>% filter(tile=='36MZD'), red = "B04", blue = "B02", green = "B08")
```
```{r plottile2, echo=TRUE}
plot(dea_s2_cube %>% filter(tile=='37MBU'), red = "B04", blue = "B02", green = "B08")
```
We can see that tile `37MBU` covers the entire extent, so we will use that tile.

```{r timeline, echo=TRUE}
dea_s2_cube <- dea_s2_cube %>% filter(tile=='37MBU')
```

## Regularising the data cube

Plotting the timeline below shows the time series looks quite regular, which should be expected given regular overpass dates of Sentinel-2.

```{r timeline2, echo=TRUE}
timeline <- sits_timeline(dea_s2_cube)

ggplot() + geom_point(aes(x=timeline, y=1))
```
To regularise the data cube, we must set the number of cores available for parallel computing. The function below returns the total number of cores available. Best performance is normally achieved by setting a value below this.

```{r cores, echo=TRUE}
detectCores()
```
The function below regularises the data cube to 20-m resolution and 10 day frequency.

This can take some time to run.

```{r regularise, echo=TRUE}
dea_s2_cube_reg <- sits_regularize(
  output_dir = tempdir(),        # directory for storing regularised output
  cube       = dea_s2_cube,      # raster cube object to regularise
  res        = 20,               # 20m spatial resolution
  period     = "P10D",           # time period
  multicores = 4                 # number of cores
)
```

Following this procedure, we can inspect the first image in the time series and note that cloud masking has been applied. 

```{r regplot, echo=TRUE}
dea_s2_cube_reg %>% 
  dplyr::filter(tile == "37MBU")  %>% 
  plot(red = "B04", blue = "B02", green = "B03")
```


```{r regtime, echo=TRUE}
timeline_reg <- sits_timeline(dea_s2_cube_reg)

ggplot() + geom_point(aes(x=timeline_reg, y=1))
```

## Calculating band indices

Now that the data cube is regularised, we can perform analysis and implement workflows like classification.

Below, NDVI is calculated and plotted.

```{r ndviplot, echo=TRUE}
reg_cube <- sits_apply(dea_s2_cube_reg,
  NDVI = (B08 - B04) / (B08 + B04),
  output_dir = tempdir()
)

plot(reg_cube, band = "NDVI", palette = "RdYlGn")
```

## Spectral mixture analysis

Mixed pixels contain a variety of spectral responses from different land use/land cover types (e.g. forest, grassland, water). If the spectral response of a pure land cover class is known (the value is called an endmember), mixture analysis provides a means to estimate the proportions of each land cover class in a pixel.

```{r ems}
em <- tibble::tribble(
  ~class,   ~B02, ~B03, ~B04, ~B8A, ~B11, ~B12,
  "forest",  200,  352,  189, 2800, 1340,  546,
  "soil",    400,  650,  700, 3600, 3500, 1800,
  "water",   700, 1100, 1400,  850,   40,   26
)
# Generate the mixture model
reg_cube <- sits_mixture_model(
  data = reg_cube,
  endmembers = em,
  multicores = 4,
  memsize = 12,
  output_dir = tempdir()
)
```

```{r forest}
plot(reg_cube, band = "FOREST", palette = "Greens")
```

```{r water}
plot(reg_cube, band = "WATER", palette = "Blues")
```

```{r soil}
plot(reg_cube, band = "SOIL", palette = "OrRd")
```

## Conclusion

We can see that the SITS package makes it easy to create and perform operations on data cubes using data from Digital Earth Africa. Next, we'll see how the package can be used for classification.